SubsidyExplorer - Step 6: Run Model

### 2020-08-11 scripts/01_data_wrangling.Rmd
updating WTO status, overseas territory info, and fao data

```{r}
# Packages
library(knitr) # Necessary to convert file to .html
library(countrycode) # Country name matching
library(here) # File path handling
library(WDI) # World Bank World Development Indicators data
library(janitor) # Column name handling
library(tidyverse) # General data wrangling
library(readxl) # Load .xlsx files

@@

# EU countries
eu_countries <- country_dependencies$iso3[country_dependencies$is_EU]
eu_country_names <- country_dependencies$WTO_name[country_dependencies$iso3 %in% eu_countries]

# EU overseas territories
eu_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 %in% eu_countries]
eu_territory_names <- sort(c(countrycode(eu_territories, "iso3c", "country.name"), "Ascension Island", "Tristan da Cunha"))

# US overseas territories:  American Samoa, Guam, Northern Marianas Islands, Puerto Rico, US Virgin Islands
us_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "USA"]
us_territory_names <- sort(countrycode(us_territories, "iso3c", "country.name"))

# Norwegian overseas territories: 
nor_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 == "NOR"]
nor_territory_names <- sort(countrycode(nor_territories, "iso3c", "country.name"))
```

Additionally, it is important to note that the European Union as a political entity is a Member of the World Trade Organization, rather then the individual states that make up the European Union. We therefore need to create a profile for the European Union as a whole.  The following `r length(eu_country_names)` states are considered to be part of the European Union for the purposes of this analysis: `r paste(eu_country_names, collapse = ", ")`.

The Netherlands, France, Denmark, and the United Kingdom also have overseas dependencies that should be considered to be a part of the European Union for statistical purposes. The following overseas dependencies are also considered to be a part of the European Union for the purpose of aggregating data in this analysis: `r paste(eu_territory_names, collapse = ", ")`.

Outside of the EU, the other dependencies we need to worry about are those that relate to the United States and Norway. The following overseas dependencies are considered to be a part of the United States for the purpose of aggregating data in this analysis: `r paste(us_territory_names, collapse = ", ")`. The following overseas dependencies are considered to be a part of Norway for the purpose of aggregating data in this analysis: `r paste(norway_territory_names, collapse = ", ")`.

@@

# Deal with the EU
sumaila_2019_dat_eu <- sumaila_2019_dat_edit %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU",
         data_type = "Aggregate") %>%
  group_by(entity, year, category, category_name, type, type_name, data_type, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity)

@@

@@

# Deal with the EU
oecd_fse_dat_eu <- oecd_fse_dat_edit %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU",
         data_type = "Aggregate") %>%
  group_by(entity, year, category, category_name, type, type_name, data_type, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, type)

@@

@@

# Save
write_csv(schuhbauer_subsidies_ssf_dat_edit, paste0(data_dir, "schuhbauer_ssf_subsidies_proportions.csv"))
write_csv(schuhbauer_subsidies_ssf_dat_edit, paste0(app_data_dir, "schuhbauer_ssf_subsidies_proportions.csv"))

@@

```{r}
# Load data using WDI package
wb_demo_dat <- WDI(country = "all",
                   indicator = c("gdp" = "NY.GDP.MKTP.CD",
                                 "gdp_ffa" = "NV.AGR.TOTL.CD",
                                 "population" = "SP.POP.TOTL"),
                                  start = 2000,
                                  end = 2018,
                                  extra = TRUE) %>%
  dplyr::filter(region != "Aggregates") %>%
  dplyr::select(iso3 = iso3c,
                year,
                gdp,
                gdp_ffa,
                population) %>%
  mutate(iso3 = as.character(iso3)) %>%
  arrange(iso3, year) %>%
  gather(variable, value, c("gdp", "gdp_ffa", "population")) %>%
  dplyr::filter(!is.na(value)) %>%
  mutate(units = case_when(variable == "population" ~ "individuals",
                           TRUE ~ paste0(year, " US$")),
         source = "World Bank WDI database")

# Deal with the EU + territories
wb_dat_eu <- wb_demo_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

# Deal with the US + territories
wb_dat_us <- wb_demo_dat %>%
  dplyr::filter(iso3 %in% c(us_territories, "USA")) %>%
  mutate(entity = "USA") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

# Add back in
wb_dat_edit <- wb_demo_dat %>%
  dplyr::filter(iso3 != "USA") %>%
  bind_rows(wb_dat_eu) %>%
  bind_rows(wb_dat_us)

# Export clean
write_csv(wb_dat_edit, paste0(data_dir, "world_bank_2020_wdi_tidy.csv"))
write_csv(wb_dat_edit, paste0(app_data_dir, "world_bank_2020_wdi_tidy.csv"))
```

@@

# Deal with EU
fao_fishers_dat_eu <- fao_fishers_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

@@

@@

# Deal with EU
teh_fte_dat_eu <- teh_fte_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

@@

@@

# Deal with the EU + territories
fao_production_dat_eu <- fao_production_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories, "EU")) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity)

# Deal with the US + territories
fao_production_dat_us <- fao_production_dat %>%
  dplyr::filter(iso3 %in% c(us_territories, "USA")) %>%
  mutate(entity = "USA-T") %>%
  group_by(entity, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  rename(iso3 = entity)

# Put them back together
fao_production_dat_edit <- fao_production_dat %>%
  dplyr::filter(iso3 != "EU") %>%
  bind_rows(fao_production_dat_eu) %>%
  bind_rows(fao_production_dat_us) %>%
  mutate(variable = "capture_production",
         units = "tonnes",
         source = "FAO (2020)")

@@

# Aggregate by ISCCAP groups and export for time series plot
fao_production_dat_edit_groups <- fao_production_dat_edit %>%
  group_by(iso3, year, isscaap_group, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T))

@@

# Look at NAs - 10,023 entries
combine_nas <- combine_catch_price_dat %>%
  dplyr::filter(is.na(exvessel_inclusive))

@@

# Calculate revenues and aggregate by ISSCAAP group
landed_value_dat_groups <- landed_value_dat %>%
  group_by(iso3, year, isscaap_group, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T))

@@

This section is a bit of a relic from when we were doing "country profiles", but there are a couple of relative metrics we want to calculate from our data sources. This is done here.

@@

>>>>

```{r}
# Packages
library(knitr) # Necessary to convert file to .html
library(countrycode) # Country name matching
library(here) # File path handling
#remotes::install_github('vincentarelbundock/WDI')
library(WDI) # World Bank World Development Indicators data
library(janitor) # Column name handling
library(tidyverse) # General data wrangling
library(readxl) # Load .xlsx files

@@

# Get all overseas dependencies that are relevant for the WTO - there are 49
territories <- country_dependencies %>%
  dplyr::filter(is_overseas_territory & is_WTO) %>%
  distinct(iso3, sovereign_iso3)

# Get all cooresponding sovereign states (we need to make sure any totals for these states include their dependencies) - there are 9
# Australia, Denmark, France, United Kingdom, Morocco, Netherlands, Norway, New Zealand, and the United States
sovereigns_with_territories <- territories %>%
  distinct(sovereign_iso3)

# EU countries
eu_countries <- country_dependencies$iso3[country_dependencies$is_EU]

# EU overseas territories
eu_territories <- country_dependencies$iso3[country_dependencies$is_overseas_territory & country_dependencies$sovereign_iso3 %in% eu_countries]
```

Additionally, it is important to note that the European Union as a political entity is a Member of the World Trade Organization, rather then the individual states that make up the European Union. We therefore need to create a profile for the European Union as a whole.  The following `r length(eu_country_names)` states are considered to be part of the European Union for the purposes of this analysis: `r paste(country_dependencies$WTO_name[country_dependencies$iso3 %in% eu_countries], collapse = ", ")`.

The Netherlands, France, Denmark, and the United Kingdom also have overseas dependencies that should be considered to be a part of the European Union for statistical purposes.

Outside of the EU, the other dependencies we need to worry about are those that relate to the Austraila, Morocco, Norway, New Zealand, and the United States.

@@

# Deal with the EU
sumaila_2019_dat_eu <- sumaila_2019_dat_edit %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU",
         data_type = "Aggregate") %>%
  group_by(entity, year, category, category_name, type, type_name, data_type, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  rename(iso3 = entity)

@@

# Test to make sure there isn't data for any other dependencies - nope we're good
sumaila_2019_dat_overseas_territory <- sumaila_2019_dat_edit %>%
  dplyr::filter(iso3 %in% territories$iso3 & !(iso3 %in% eu_territories))

@@

# Deal with the EU
oecd_fse_dat_eu <- oecd_fse_dat_edit %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU",
         data_type = "Aggregate") %>%
  group_by(entity, year, category, category_name, type, type_name, data_type, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, type)

@@

# Test to make sure there isn't data for any other dependencies - nope we're good
oecd_fse_dat_overseas_territory <- oecd_fse_dat_edit %>%
  dplyr::filter(iso3 %in% territories$iso3 & !(iso3 %in% eu_territories))

@@

# Test to make sure there isn't data for any other dependencies - nope we're good
schuhbauer_subsidies_ssf_dat_overseas_territory <- schuhbauer_subsidies_ssf_dat_edit %>%
  dplyr::filter(iso3 %in% territories$iso3 & !(iso3 %in% eu_territories))

# Save
write_csv(schuhbauer_subsidies_ssf_dat_edit, paste0(data_dir, "schuhbauer_ssf_subsidies_proportions.csv"))
write_csv(schuhbauer_subsidies_ssf_dat_edit, paste0(app_data_dir, "schuhbauer_ssf_subsidies_proportions.csv"))
```

@@

```{r}
# Load data using WDI package'
# wb_demo_dat <- WDI(country = "all",
#                    indicator = c("gdp" = "NY.GDP.MKTP.CD",
#                                  "gdp_ffa" = "NV.AGR.TOTL.CD",
#                                  "population" = "SP.POP.TOTL"),
#                                   start = 2000,
#                                   end = 2018,
#                                   extra = TRUE) %>%
#   dplyr::filter(region != "Aggregates") %>%
#   dplyr::select(iso3 = iso3c,
#                 year,
#                 gdp,
#                 gdp_ffa,
#                 population) %>%
#   mutate(iso3 = as.character(iso3)) %>%
#   arrange(iso3, year) %>%
#   gather(variable, value, c("gdp", "gdp_ffa", "population")) %>%
#   dplyr::filter(!is.na(value)) %>%
#   mutate(units = case_when(variable == "population" ~ "individuals",
#                            TRUE ~ paste0(year, " US$")),
#          source = "World Bank WDI database")
# 
# # Deal with the EU + territories
# wb_dat_eu <- wb_demo_dat %>%
#   dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
#   mutate(entity = "EU") %>%
#   group_by(entity, year, variable, units, source) %>%
#   summarize(value = sum(value, na.rm = T)) %>%
#   rename(iso3 = entity) %>%
#   arrange(iso3, year, variable)
# 
# # Test to make sure there isn't data for any other dependencies - will need to add this once I can figure out why the WDI API isn't working
# wb_dat_overseas_territory <- wb_demo_dat %>%
#   dplyr::filter(iso3 %in% territories$iso3 & !(iso3 %in% eu_territories))
# # Deal with the US + territories
# wb_dat_us <- wb_demo_dat %>%
#   dplyr::filter(iso3 %in% c(us_territories, "USA")) %>%
#   mutate(entity = "USA") %>%
#   group_by(entity, year, variable, units, source) %>%
#   summarize(value = sum(value, na.rm = T)) %>%
#   rename(iso3 = entity) %>%
#   arrange(iso3, year, variable)
# 
# # Add back in
# wb_dat_edit <- wb_demo_dat %>%
#   dplyr::filter(iso3 != "USA") %>%
#   bind_rows(wb_dat_eu) %>%
#   bind_rows(wb_dat_us)
# 
# # Export clean
# write_csv(wb_dat_edit, paste0(data_dir, "world_bank_2020_wdi_tidy.csv"))
# write_csv(wb_dat_edit, paste0(app_data_dir, "world_bank_2020_wdi_tidy.csv"))

# Load in
wb_dat_edit <- read_csv(paste0(data_dir, "world_bank_2020_wdi_tidy.csv"))
```

@@

# Deal with EU
fao_fishers_dat_eu <- fao_fishers_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

@@

# # Test to make sure there isn't data for any other dependencies - nope we're good
fao_fishers_dat_overseas_territory <- fao_fishers_dat %>%
  dplyr::filter(iso3 %in% territories$iso3 & !(iso3 %in% eu_territories))

@@

# Deal with EU
teh_fte_dat_eu <- teh_fte_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories)) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  rename(iso3 = entity) %>%
  arrange(iso3, year, variable)

@@

# # Test to make sure there isn't data for any other dependencies - nope we're good
teh_fte_dat_overseas_territory <- teh_fte_dat_edit %>%
  dplyr::filter(iso3 %in% territories$iso3 & !(iso3 %in% eu_territories))

@@

# Deal with the EU + territories
fao_production_dat_eu <- fao_production_dat %>%
  dplyr::filter(iso3 %in% c(eu_countries, eu_territories, "EU")) %>%
  mutate(entity = "EU") %>%
  group_by(entity, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  rename(iso3 = entity)

# Deal with all territories 
fao_production_dat_overseas_territories <- fao_production_dat %>%
  dplyr::filter(iso3 %in% territories$iso3 | iso3 %in% territories$sovereign_iso3) %>%
  left_join(territories, by = c("iso3")) %>%
  mutate(sovereign_iso3 = case_when(is.na(sovereign_iso3) ~ iso3,
                                    TRUE ~ sovereign_iso3)) %>%
  mutate(entity = paste0(sovereign_iso3, "-T")) %>%
  group_by(entity, year, fishing_area, species, asfis_species, scientific_name, yearbook_group, isscaap_group) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  rename(iso3 = entity)

# Put them back together
fao_production_dat_edit <- fao_production_dat %>%
  dplyr::filter(iso3 != "EU") %>%
  bind_rows(fao_production_dat_eu) %>%
  bind_rows(fao_production_dat_overseas_territories) %>%
  mutate(variable = "capture_production",
         units = "tonnes",
         source = "FAO (2020)")

@@

# Aggregate by ISCCAP groups and export for time series plot
fao_production_dat_edit_groups <- fao_production_dat_edit %>%
  group_by(iso3, year, isscaap_group, variable, units, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  spread(year, value) %>% # adding empty rows to make the plotting work better
  gather(year, value, as.character(c(2000:2018))) %>% # second part of that process
  mutate(year = as.numeric(year))
fao_production_dat_edit_groups$value[is.na(fao_production_dat_edit_groups$value)] <- 0 # for plotting purposes - we don't get gaps on area plot this way

@@

# Look at NAs - 11,553 entries
combine_nas <- combine_catch_price_dat %>%
  dplyr::filter(is.na(exvessel_inclusive))

combine_nas_groups <- unique(combine_nas$isscaap_group)

@@

# Calculate revenues and aggregate by ISSCAAP group
landed_value_dat_groups <- landed_value_dat %>%
  group_by(iso3, year, isscaap_group, variable, source) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  spread(year, value) %>% # adding empty rows to make the plotting work better
  gather(year, value, as.character(c(2000:2018))) %>% # second part of that process
  mutate(year = as.numeric(year)) %>%
  mutate(units = paste0(year, " US$"))

@@

This section is a bit of a relic from when we were doing "country profiles", but there are a couple of relative subsidy metrics we want to calculate from our data sources. This is done here.

@@

# Section 6 - Cap/Tier Lookup Tables

Many of the capping proposals utilize some metric to sort Members into tiers. This is often done based upon the average over some period. In order to allow this to be interactive later on, we need to craft a lookup table with all possible date ranges.

```{r}
# Capture production lookup
capture_production_lookup <- fao_production_dat_edit_groups %>%
  group_by(iso3, year) %>%
  summarize(capture_production = sum(value, na.rm = T)) %>%
  ungroup() %>%
  spread(year, capture_production) %>%
  mutate(variable = "capture_production") %>%
  left_join(country_dependencies %>% dplyr::select(iso3, is_WTO), by = "iso3") %>%
  mutate(is_WTO = case_when(!is.na(is_WTO) ~ is_WTO,
                            iso3 %in% unique(paste0(territories$sovereign_iso3, "-T")) ~ TRUE,
                            iso3 == "EU" ~ TRUE,
                            TRUE ~ FALSE))

# Landed value
landed_value_lookup <- landed_value_dat_groups %>%
  group_by(iso3, year) %>%
  summarize(landed_value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  spread(year, landed_value) %>%
  mutate(variable = "landed_value") %>%
  left_join(country_dependencies %>% dplyr::select(iso3, is_WTO), by = "iso3") %>%
  mutate(is_WTO = case_when(!is.na(is_WTO) ~ is_WTO,
                            iso3 %in% unique(paste0(territories$sovereign_iso3, "-T")) ~ TRUE,
                            iso3 == "EU" ~ TRUE,
                            TRUE ~ FALSE))

# Combine 
cap_tier_lookup <- capture_production_lookup %>%
  bind_rows(landed_value_lookup)

write_csv(cap_tier_lookup, paste0(data_dir, "cap_tier_lookup_table.csv"))
write_csv(cap_tier_lookup, paste0(app_data_dir, "cap_tier_lookup_table.csv"))

### 2020-08-11 scripts/06_run_model.Rmd
updating WTO status, overseas territory info, and fao data

# Load country dependencies helper file
country_dependencies <- read_csv(here::here("data", "lookup-tables", "country_dependencies.csv")) %>%
  mutate(territory_is_WTO = case_when(WTO_status %in% c("Member", "Observer") ~ T,
                            TRUE ~ F))

sovereign_statuses <- country_dependencies %>%
  distinct(sovereign_iso3, WTO_status) %>%
  mutate(sovereign_is_WTO = case_when(WTO_status %in% c("Member", "Observer") ~ T,
                            TRUE ~ F))

wto_lookup <- country_dependencies %>%
  left_join(sovereign_statuses, by = c("sovereign_iso3", "WTO_status")) %>%
  mutate(is_WTO = case_when(territory_is_WTO ~ T,
                            sovereign_is_WTO ~ T,
                            TRUE ~ F)) %>%
  dplyr::select(iso3, is_WTO)
```

@@

```{r pressure, echo=FALSE}
# Assign fleets to vessels
global_fleet_vessels <- vessel_dat %>%
  left_join(wto_lookup, by = c("flag_iso3" = "iso3")) %>%
  mutate(fleet = case_when(is_WTO & fmi_best >= managed_cutoff & bad_subs > 0 ~ "affected_managed",
                           is_WTO & fmi_best < managed_cutoff & bad_subs > 0 ~ "affected_oa",
                           (!is_WTO & fmi_best >= managed_cutoff) ~ "unaffected_managed",
                           (!is_WTO & fmi_best < managed_cutoff) ~ "unaffected_oa",
                           TRUE ~ "unaffected_oa"))
```

>>>>

# Load country dependencies helper file
country_dependencies <- read_csv(here::here("data", "lookup-tables", "country_dependencies.csv"))
```

@@

```{r pressure, echo=FALSE}
# Assign fleets to vessels
global_fleet_vessels <- vessel_dat %>%
  left_join(country_dependencies %>% dplyr::select(iso3, is_WTO), by = c("flag_iso3" = "iso3")) %>%
  mutate(fleet = case_when(is_WTO & fmi_best >= managed_cutoff & bad_subs > 0 ~ "affected_managed",
                           is_WTO & fmi_best < managed_cutoff & bad_subs > 0 ~ "affected_oa",
                           (!is_WTO & fmi_best >= managed_cutoff) ~ "unaffected_managed",
                           (!is_WTO & fmi_best < managed_cutoff) ~ "unaffected_oa",
                           TRUE ~ "unaffected_oa"))
```
